A biometric analysis of infant mortality and temperature, northern Sweden 1895-1950

### 2022-02-09 getdata4.Rmd
new births.rds

```{r fromobs}
bobs <- obs[obs$starttyp == 2 & obs$id %in% per$id, ]
indx <- match(per$id, bobs$id)
per$socStatus <- bobs$socStatus[indx]
per$socBranch <- bobs$socBranch[indx]
per$hisclass <- bobs$hisclass[indx]
per$nofrs <- bobs$nofrs[indx]
per <- per[!is.na(per$nofrs), ]
per$subreg <- factor(per$nofrs)
levels(per$subreg) <- c(rep("ume", 2), 
                        rep("bjur", 4), 
                        rep("sten", 3))
##per <- per[(per$subreg != "sten"), ] # NOTE: Now Stensele station is included
per$subreg <- factor(per$subreg)
per$enter <- 0
per$sex <- factor(per$kon, labels = c("boy", "girl"))
per$birthdate <- toTime(per$birthdate)
per$illeg <- as.numeric(per$ab == 2)
per$parity <- per$paritet_g
per$paritet_g <- per$paritet_k <- per$doddat <- per$frsbosdat <- NULL
per$fid <- per$fboort <- per$ab <- per$fyrkfd <- per$fyrkind <- NULL
per$region <- per$kon <- per$fodhfrs <- per$frsbobdat <- per$frsbostyp <- NULL
per$fbonmn <- per$fboind <- per$fyrkrel <- per$fyrketxt <- NULL
per$foddat <- as.Date(as.character(per$foddat), format = "%Y%m%d")
## NEW:
per$socBranch <- as.character(per$socBranch)
per$socBranch[is.na(per$socBranch)] <- "none"
per$socBranch <- factor(per$socBranch)
per$socStatus <- as.character(per$socStatus)
per$socStatus[is.na(per$socStatus)] <- "low"
per$socStatus <- factor(per$socStatus)
per <- per[, c("id", "sex", "birthdate", "mid", "foddat", "parity", "illeg", 
               "subreg", "socBranch", "socStatus", "hisclass", "enter", "exit", "event")]
#saveRDS(per, file = "data/infdat3.rds")
```

>>>>

```{r fromobs}
bobs <- obs[obs$starttyp == 2 & obs$id %in% per$id, ]
indx <- match(per$id, bobs$id)
per$socStatus <- bobs$socStatus[indx]
per$socBranch <- bobs$socBranch[indx]
per$hisclass <- bobs$hisclass[indx]
per$nofrs <- bobs$nofrs[indx]
per <- per[!is.na(per$nofrs), ]
per$subreg <- factor(per$nofrs)
levels(per$subreg) <- c(rep("ume", 2), 
                        rep("bjur", 4), 
                        rep("sten", 3))
##per <- per[(per$subreg != "sten"), ] # NOTE: Now Stensele station is included
per$subreg <- factor(per$subreg)
per$enter <- 0
per$sex <- factor(per$kon, labels = c("boy", "girl"))
per$birthdate <- toTime(per$birthdate)
per$illeg <- as.numeric(per$ab == 2)
per$parity <- per$paritet_g
per$paritet_g <- per$paritet_k <- per$doddat <- per$frsbosdat <- NULL
per$fid <- per$fboort <- per$ab <- per$fyrkfd <- per$fyrkind <- NULL
per$region <- per$kon <- per$fodhfrs <- per$frsbobdat <- per$frsbostyp <- NULL
per$fbonmn <- per$fboind <- per$fyrkrel <- per$fyrketxt <- NULL
per$foddat <- as.Date(as.character(per$foddat), format = "%Y%m%d")
## NEW:
per$socBranch <- as.character(per$socBranch)
per$socBranch[is.na(per$socBranch)] <- "none"
per$socBranch <- factor(per$socBranch)
per$socStatus <- as.character(per$socStatus)
per$socStatus[is.na(per$socStatus)] <- "low"
per$socStatus <- factor(per$socStatus)
per <- per[, c("id", "sex", "birthdate", "mid", "foddat", "parity", "illeg", 
               "subreg", "socBranch", "socStatus", "hisclass", "enter", "exit", "event")]
saveRDS(per, file = "data/births4.rds")
```

### 2022-02-09 paper4.Rmd
new births.rds

---
title: "A biometric analysis of infant mortality and temperature, northern Sweden 1895-1950"
author: "Göran Broström and Tommy Bengtsson"
date: "`r Sys.time()`"
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 3
  bookdown::word_document2:
    toc: no
    toc_depth: 2
  bookdown::html_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
citation_package: natbib
classoption: titlepage
bibliography: bio.bib
titlepage: yes
biblio-style: apalike
documentclass: article
fontsize: 11pt
header-includes:
- \usepackage[utf8]{inputenc}
- \usepackage{graphicx}
- \usepackage{a4wide}
- \usepackage[tablesonly]{endfloat}
- \renewcommand{\efloatseparator}{\mbox{}}
abstract: "The effect of extreme temperatures on infant mortality in the Umeå and Skellefteå regions 1895-1950 is studied in a biometric analysis setting. More precisely, the effect of climate and weather, measured by temperature, average and extremes, on infant mortality is investigated. It turns out that climate (average) is more important than weather (extremes), low average temperatures are more important than temporary dips in temperature."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.path='figs/',fig.height=3, comment=NA, cache= TRUE) ## NOTE!!
knitr::opts_knit$set(global.par = TRUE)
##
library(knitr)
library(eha)
```

@@

```{r readdata}
births <- readRDS("data/births.rds")
infdat  <- readRDS("data/infdat4.rds")
```
# Introduction

@@

>>>>

---
title: "A biometric analysis of infant mortality and temperature, northern Sweden 1895-1950"
author: "Göran Broström and Tommy Bengtsson"
date: "`r Sys.time()`"
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 3
  bookdown::word_document2:
    toc: no
    toc_depth: 2
  bookdown::html_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
citation_package: natbib
classoption: titlepage
bibliography: bio.bib
titlepage: yes
biblio-style: apalike
documentclass: article
fontsize: 11pt
header-includes:
- \usepackage[utf8]{inputenc}
- \usepackage{graphicx}
- \usepackage{a4wide}
- \usepackage[tablesonly]{endfloat}
- \renewcommand{\efloatseparator}{\mbox{}}
abstract: "The effect of extreme temperatures on infant mortality in the Umeå and Skellefteå regions 1895-1950 is studied in a biometric analysis setting. More precisely, the effect of climate and weather, measured by temperature, average and extremes, on infant mortality is investigated. It turns out that climate (average) is more important than weather (extremes), low average temperatures are more important than temporary dips in temperature, but effects are different in neonatal and postneonatal settings."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.path='figs/',fig.height=3, comment=NA, cache= TRUE) ## NOTE!!
knitr::opts_knit$set(global.par = TRUE)
##
library(knitr)
library(eha)
starttid <- proc.time()
```

@@

```{r readdata}
births <- readRDS("data/births4.rds")
infdat  <- readRDS("data/infdat4.rds")
```
# Introduction

@@

Used processing time (seconds):

```{r tidtagning}
sluttid <- Sys.time()
proc.time() - starttid
```

Done: `r sluttid`.
